---
title: "What Predicts Students' Final Scores?"
author: "Muktadir Somio"
date: today
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(broom)

dat <- read_csv("stdnt_perf_clean.csv")

fit <- lm(final_score ~ study_hours_week + attendance_pct + prior_gpa +
            sleep_hours + part_time_job, data = dat)
est <- tidy(fit, conf.int = TRUE)
study_est <- est |> filter(term == "study_hours_week")
job_est   <- est |> filter(term == "part_time_jobYes")
r2 <- glance(fit)$r.squared
```

## Summary

We studied `r nrow(dat)` students after cleaning the raw file.
Each extra weekly study hour is linked with about
`r round(study_est$estimate, 2)` more points on the final exam
(95% CI `r round(study_est$conf.low, 2)` to `r round(study_est$conf.high, 2)`),
holding the other factors fixed. Students with a part-time job score about
`r round(abs(job_est$estimate), 1)` points lower. Together, the model explains
`r round(100 * r2)`% of the variation in final scores.

## Data

```{r}
#| label: tbl-data
dat |>
  summarise(Students = n(),
            `Mean final score` = round(mean(final_score), 1),
            `Pass rate` = scales::percent(mean(passed), accuracy = 1)) |>
  knitr::kable()
```

## Main finding

```{r}
#| label: fig-study
#| fig-cap: "Study hours and final score (line = linear fit)."
ggplot(dat, aes(study_hours_week, final_score)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Study hours per week", y = "Final score")
```

## Model results

```{r}
#| label: tbl-model
est |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  select(Term = term, Estimate = estimate, `Std. error` = std.error,
         `p-value` = p.value, Lower = conf.low, Upper = conf.high) |>
  knitr::kable()
```

## Limits

- This is observational data, so the results show association, not cause.
- Rows with missing values were left out of the model.
- The model explains part of the variation, not all of it.